---
title: "Continental Shelf Depth"
format: html
---

The depth of the edge of the continental shelf (the shelf break) has multiple definitions in the literature, including 133 and 167m ([Hay 2016](https://doi.org/10.1007/978-94-007-6238-1_179)),  200m ([Harris et al. 2014](https://bluehabitats.org/?page_id=1660)), and 300m ([Watling et al. 2013](https://doi.org/10.1016/j.pocean.2012.11.003)). 

Here, I calculate the global average continental shelf break depth, using the edge of the continental shelf as defined by Harris et al. ([2014](https://doi.org/10.1007/978-94-007-6238-1_179)), and extracting depth data from the [GEBCO 2024 sub-ice bathymetry](https://www.gebco.net/).

```{r}
#| include: false

library(terra)
```

```{r}
shelf <- vect("../data/Shelf.shp")

plet(shelf[,1])
```

Try getting only the longest lines of each shelf polygon, but this doesn't work comprehensively because some of the shelf breaks touch the coastline creating a single polygon that can't be split into separate lines.

```{r}
nrow(shelf)

shelf_list <- lapply(1:nrow(shelf), FUN = function(x) shelf[x, 1] |> 
                       as.lines() |> 
                       disagg())

#how many elements have only 1 line?

no_features <- sapply(shelf_list, nrow)

table(no_features)
                       
for (i in 1:length(shelf_list)) {
  if(nrow(shelf_list[[i]])>1){
    shelf_list[[i]]$length <- perim(shelf_list[[i]]) #add column with line lengths
    
    row_index_longest <- which(shelf_list[[i]]$length == max(shelf_list[[i]]$length)) #find which row contains the line with the longest length
    
    shelf_list[[i]] <- shelf_list[[i]][row_index_longest, 1]
  } 
}

shelf_lines <- do.call(rbind, shelf_list)
```

Try erasing land using rnaturalearth: this doesn't work either because I need a larger buffer than I've applied to erase all land and then I will be erasing some of the shelf breaks!

```{r}
land <- rnaturalearth::ne_countries(scale = 10, returnclass = "sv") |> 
  aggregate()

land_buffered <- buffer(land, width = 100)

shelf_no_land <- erase(shelf, land_buffered)
```

